---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

#Upload data
```{r}
D1 <- read.csv("online.data.csv", header = TRUE)
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)

#Tidy data
library(dplyr)
library(tidyr)
D1 <- select(D1, -id)
D1$level.up <- ifelse(D1$level.up=="yes",1,0)
D2 <- gather(D1, key=variable, value=value)

#Histograms with "facet"
library(ggplot2)
hist <- ggplot(D2, aes(x=value)) + geom_histogram(binwidth=0.05) + facet_wrap(~variable,scales="free")
hist
##ggplot(D2, aes(x=value)) + geom_histogram(binwidth=0.1) + facet_grid(variable ~.,scales="free")

#Then visualize the relationships between variables
pairs(D1, labels=colnames(D1), main="Relationships Between Variables")

#Correlation Plot
library(corrplot)
COR <- cor(D1)
corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

#Try to capture an intution about the data and the relationships
## "post.test.score" and "messages" may have linear relationship
## "level.up" may have relationship with "post.test.score", "messages", and "av.assignment.score"

```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
library(rpart)
c.tree1 <- rpart(level.up~post.test.score+messages+av.assignment.score, method="class", data=D1)

#Plot and generate a CP table for your tree 
post(c.tree1, file = "prediction1.ps", title = "Level Up 1")
printcp(c.tree1)

#Generate a probability value that represents the probability that a student levels up based your classification tree 

D1$pred1 <- predict(c.tree1, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.

#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.

library(ROCR)

#Plot the curve
pred.detail1 <- prediction(D1$pred1, D1$level.up) 
plot(performance(pred.detail1, "tpr", "fpr"))
#plot(performance(pred.detail1, "tpr", "fpr"),colorize=TRUE)
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail1,"auc"), "y.values"))#Unlist liberates the AUC value from the "performance" object created by ROCR
## AUC1=1

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?

c.tree2 <- rpart(level.up~pre.test.score+forum.posts, method="class", data=D1)
post(c.tree2, file = "prediction2.ps", title = "Level Up 2")
printcp(c.tree2)

D1$pred2 <- predict(c.tree2, type = "prob")[,2]
pred.detail2 <- prediction(D1$pred2, D1$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)
unlist(slot(performance(pred.detail2,"auc"), "y.values"))
## AUC2=0.82

## Model1 is better. Because it has larger AUC than model2
```
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

#threshold1: 0
D1$threshold.pred1 <- ifelse(D1$pred1 > 0.8, 1, 0) #true positive + false positive

#Now generate three diagnostics:

D1$TPTN <- ifelse(D1$level.up==D1$threshold.pred1,1,0) #true positive + true negative
accuracy.model1 <- sum(D1$TPTN) / nrow(D1)

D1$TP <- ifelse(D1$threshold.pred1==1 & D1$level.up==1,1,0) #true positive
precision.model1 <- sum(D1$TP) / sum(D1$threshold.pred1)

D1$FN <- ifelse(D1$threshold.pred1==0 & D1$level.up==1,1,0) #false negative
recall.model1 <- sum(D1$TP) / (sum(D1$TP)+sum(D1$FN))

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(D1$level.up, D1$threshold.pred1)
table1
#accuracy.model1=(400+600)/1000
#precision.model1=400/400
#recall.model1=400/400

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1)/kappa(matrix1, exact = TRUE)

#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

#threshold2: 0.5
D1$threshold.pred2 <- ifelse(D1$pred1 > 0.5, 1, 0) #true positive + false positive
table2 <- table(D1$level.up, D1$threshold.pred2)
table2
accuracy.model2=(400+600)/1000
precision.model2=400/400
recall.model2=400/400

matrix2 <- as.matrix(table2)
kappa(matrix2)/kappa(matrix2, exact = TRUE)

##Conclusion: The changes of threshold in the "perfect" model did not influence the diagnostics results.
##The clear cutoff in the "perfect" model indicates that changing the threshold above a very low number will result in many more false positives with no improvement in true positives.

```


